---
output: html_document
editor_options: 
  chunk_output_type: console
---
RED WINE EDA CONNOR ZARKOWSKI
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(GGally)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wine = read.csv("wineQualityReds.csv")
```

The following dataset contains information about a sample of red wines. Our \
goal is to explore the data to see how it is structured, and analyze \
relationships between the variables. Primarily I want to see if there are any\
strong relationships between quality and the other variables. 

```{r echo=FALSE}
str(wine)
summary(wine)
```

Looking at the structure and summary of the data we can get an idea of what \
the dataset looks like and the range of values for each variable. Next we'll \
look more closely at individual variables.

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}
qplot(alcohol, data= wine, geom= "histogram", color =I('black'), 
      fill = I('#008080'), bins= 40)
```

```{r echo=FALSE}
qplot(density, data= wine, geom= "histogram", color =I('black'), 
      fill = I('#008080'), bins= 30)
```

```{r echo=FALSE}
qplot(pH, data= wine, geom= "histogram", color =I('black'), 
      fill = I('#008080'), bins= 30)
```

```{r echo=FALSE}
qplot(quality, data= wine, geom= "histogram", color =I('black'), 
      fill = I('#008080'), bins= 16)
```

```{r echo=FALSE}
qplot(citric.acid, data= wine, geom= "histogram", color =I('black'), 
      fill = I('#008080'), bins= 30)
```

```{r echo=FALSE}
qplot(fixed.acidity, data= wine, geom= "histogram", color =I('black'), 
      fill = I('#008080'), bins= 30)
```

```{r echo= FALSE}
qplot(volatile.acidity, data= wine, geom= "histogram", color =I('black'), 
      fill = I('#008080'), bins= 30)
```

The distributions of the variables of interest either follow a normal or skewed\
distribution. The average alcohol percentage is 10.4. PH falls between a range \
of 2.7 and 4. Quality ranges from a rating of 3 to 8. 

# Univariate Analysis

### What is the structure of your dataset?

The dataset contains 1599 wines, with each observation having 13 \
variables corresponding to it. 

### What is/are the main feature(s) of interest in your dataset?

The main features are alcohol, density, fixed and volatile acidity, and pH. \
These are metricsI've heard people talk about when tasting wine.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

When looking at fixed and volatile acidity it will be helpful to see the \
relationship they have with citric acid and pH. 

### Did you create any new variables from existing variables in the dataset?

No new variables were created.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

There were no unusual distributions, they were all either normal or skewed to \
one side. I did not perform any operations on the data. 


# Bivariate Plots Section

```{r echo=FALSE, fig.width=15, fig.height=9}
ggpairs(wine)
```

Running ggpairs, quality was most strongly correlated with alcohol and \
volatile acidity. Fixed acidity was strongly correlated with density, pH and \
citric acid.

```{r echo=FALSE, Bivariate_Plots}
ggplot(aes(x= quality, y= alcohol), data= wine) + geom_jitter()
```

There is a positive correlation between alcohol and quality.

```{r echo=FALSE}
ggplot(aes(x= density, y= alcohol), data= wine) + geom_jitter()
```

There is a moderate correlation between alcohol and density. 

```{r echo=FALSE}
ggplot(aes(x= quality, y= volatile.acidity), data= wine) + geom_jitter()
```

There is a negative correlation between volatile acidity and quality, it also \
appears that the range of values for volatile.acidity decrease as quality \
increases. 

```{r echo=FALSE}
ggplot(aes(x= volatile.acidity, y= citric.acid), data= wine) + geom_jitter()
```

There is a negative correlation between the two variables, the higher the \
volatile acidity, the lower amount of citric acid.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Alcohol and volatile acidity had the strongest correlation to quality. Density \
had a strong relatinship to alcohol, and volatile and citric acid were also \
strongly correlated. I had thought sulphates would be related to quality, \
however there weren't any notable relationships. 

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

There appeared to be a relationship between all the varibles having to do with \
acid (pH, fixed.acidity, volatile.acidity, and citric.acid), it will be \
interesting to do a multivariate analysis between them.

### What was the strongest relationship you found?

The strongest relationship was between fixed acidity and pH. This was an \
expected relationship since wine gets most of it's acidity from the grapes\
themselves, and pH is a measure of overall acidity. 

# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x= quality, y= alcohol, color= density), data= wine) + geom_jitter()
```

We can see here that although density and quality aren't strongly correlated, \
there appears to be a trend where as alcohol and quality rise, density also \
seems to decrease.  

```{r echo=FALSE}
ggplot(aes(x=alcohol, y=density, color=fixed.acidity), data= wine) + 
  geom_jitter()
```

Density had significant correlations with both fixed acidity and alcohol, so I \
wanted to visualize the three variables together. We can see that as alcohol \
increases, density and fixed acidity decrease. 

```{r echo=FALSE}
ggplot(aes(x= quality, y= volatile.acidity, color= citric.acid), data= wine) +
  geom_jitter()
```

Volatile acidity had the second strongest correlation to quality, and was \
negatively correlated with citric acid. In this plot we can see both the \
negative correlation between volatile acidity and quality, and the negative \
correlation between volatile acidity and citric acid. 

```{r echo=FALSE}
ggplot(aes(x= volatile.acidity, y= citric.acid), data= wine) + geom_jitter() +
  facet_wrap(~quality)
```

Here I wanted to see if the correlation between volatile acidity and citric \
acid changed as quality increased, it doesn't appear to make a difference. 

```{r echo=FALSE}
p1<-ggplot(aes(x=citric.acid, y= fixed.acidity, color= pH), data= wine) + 
  geom_jitter()
p2<-ggplot(aes(x=citric.acid, y= volatile.acidity, color= pH), data= wine) + 
  geom_jitter()
grid.arrange(p1,p2,ncol=2)
```

Fixed acidity, volatile acidity, and citric acid all affect the pH of the wine, \
so I wanted to make a visualization of all four variables to help illustrate \
trends between them. We can see that citric acid and pH are positively \
correlated, while fixed and volatile acidty have inverse relationships to \
citric acid. 

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?


### Were there any interesting or surprising interactions between features?

I was surprised to see the inverse correlation between fixed and volatile \
acidity and citric acid. I also suspected that as wine quality got higher there \
would be less use of citirc acid as an additive, but it seems that higher \
quality wines still use it. 

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x= quality, y= alcohol, color= density), data= wine) + 
  geom_jitter() + ggtitle('Quality, Alcohol, and Density')
```

This plot shows the relationship between alcohol and quality, as well as the \
relationship between alcohol and density. As the quality of wine increases \
there is a positive correlation with the percentage of alcohol and an overall \
decrease in the wine's density.


### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(aes(x= volatile.acidity, y= citric.acid), data= wine) + 
  geom_jitter(alpha=0.4) + facet_wrap(~quality) + 
  ggtitle('Correlation of Volatile Acidity and Citric Acid by Quality')
```

### Description Two
Citric acid can often be added after wine has fermented to give it a fresher \
taste, I wanted to see if citric acid use decreased as the quality of the \
wine increased, but it seems that the average amount used stays consistent.

### Plot Three
```{r echo=FALSE, Plot_Three}
p1<-ggplot(aes(x=citric.acid, y= fixed.acidity, color= pH), data= wine) + 
  geom_jitter()
p2<-ggplot(aes(x=citric.acid, y= volatile.acidity, color= pH), data= wine) + 
  geom_jitter()
grid.arrange(p1,p2,ncol=2, top= 'Acidic Variables and pH')
```

### Description Three
While volatile acidity was the only variable with a significant correlation to \
quality, all of the variables having to do with acid had relationships with \
each other. This plot shows that higher instances of fixed acidity, volatile \
acidity, or citric acid correspond to a lower, more acidic pH. 

# Reflection

It appears that there is not a single variable that can predict the quality of\
a wine, it is the combination of many variables. Alcohol and volatile acidity \
are correlated with quality, but they also have correlations with other \
variables in the dataset. I have almost no background in wine, so I struggled \
in the beginning to understand what each of the variables was measuring. I \
would like to know more about the sample of wines, where it was from and what \
the ages were of each wine. In the future further analysis could be done with \
the sulpher-related variables, I have heard the term when people talk about \
wine, but did not find any meaningful relationships during my analysis. Also a \
linear model could be built to predict a wine's quality. 